### 2023-06-12 python_excel_ite.tex
working on intro
working on intro sections

% TODO - how should pandas appear in text
Python has continued to make inroads to analytical territory that was formerly dominated by Excel. The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst. However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models. The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful. For some time I had been thinking about exploring the general problem of doing spreadsheet type modeling with Python.

@@

@@

PCDA ends with intro to sklearn and AAP begins with review of sklearn and some more ML content - logistic regression with regularization

What kind of Python background do students need to have in order to be ready for this EwP module? This module does presume a basic familiarity with Python fundamentals for data analysis such as:

@@

The EwP module could easily be used as part of a semester long Python based analytics course. At our institution, this module is part of a business course entitled "Advanced Analytics with Python" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}). The students taking this course have already taken my "Practical Computing for Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics. Some Linux basics are also part of the PCDA course. The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed. We will describe the AAP course web in more detail below in the context of the EwP module.

@@

Combination of Jupyter notebooks, Python scripts, videos walking through the notebooks and scripts, all publicly accessible from submodule specific course web pages. GIVE URL. Some background on Jupyter notebooks for analytics. Questions and coding challenges sprinkled throughout the notebooks - answers at bottom. While AAP is online asynch, this same approach has been used very successfully in a computer teaching lab for the PCDA course. The screencasts are highly valued by the students as they can rewatch things as needed. Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.

@@

The first notebook in this submodule starts by explaining what Python package are and their role in sharing code. We then revisit a concept from earlier in the course - creating a good project folder structure. For this we use what is known as a \textit{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts. Their are a few different folder layouts used in Python projects and there is a bit of discussion on the folder stucture we will use and resources given for exploring this issue further. Students are reminded of the importance of version control and we initialize a git repository for our budding package. The basics of version control were already covered in the first week of the course. My experience is that most students have little exposure to simple software project management concepts such as project folder structures and version control. This can lead to a jumble of files with incoherent names used as a proxy for version control - whatif_v1.py, whatif_v2.py, whatif_final.py, whatif_finalfinal.py, ...

@@

While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of copy-paste operations is something that many Excel users will admit to doing. We also use this example to introduce the newish pathlib Python library for working with file paths. After getting familiar with pathlib, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.

>>>>

% TODO - how should pandas appear in text
Python has continued to make inroads to analytical territory that was formerly dominated by Excel. The widely used Python data analysis library, Pandas, was created by Wes McKinney while working as a quantitative financial analyst. However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models. The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful. But, it got me to thinking. What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models? What would those look like in Python?

@@

Created the ``Excel with Python'' (EwP) module as part

@@

At our institution, this module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}). The students taking this course have already taken my ``Practical Computing for Business Analytics'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics. Basic Linux skills are also part of the PCDA course. The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.  The EwP module could certainly be used as part of a semester long Python based analytics course. The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed. We will describe the AAP course web in more detail below in the context of the EwP module.

What kind of Python background do students need to have in order to be ready for the EwP module? This module presumes a basic familiarity with Python fundamentals for data analysis such as:

@@

@@

All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages. GIVE URL. Some background on Jupyter notebooks for analytics. Questions and coding challenges sprinkled throughout the notebooks - answers at bottom. While AAP is online asynch, this same approach has been used very successfully in a computer teaching lab for the PCDA course. The screencasts are highly valued by the students as they can rewatch things as needed. Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.

@@

The first notebook in this submodule starts by explaining what Python package are and their role in sharing code. We then revisit a concept from earlier in the course - creating a good project folder structure. For this we use what is known as a \textit{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{cookiecutter}. Their are a few different folder layouts used in Python projects and there is a bit of discussion on the folder stucture we will use and resources given for exploring this issue further. Students are reminded of the importance of version control and we initialize a git repository for our budding package. The basics of version control were already covered in the first week of the course. My experience is that most students have little exposure to simple software project management concepts such as project folder structures and version control. This can lead to a jumble of files with incoherent names used as a proxy for version control - whatif_v1.py, whatif_v2.py, whatif_final.py, whatif_finalfinal.py, ...

@@

While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of copy-paste operations is something that many Excel users will admit to doing. We also use this example to introduce the relatively new pathlib library for working with file paths. After getting familiar with pathlib, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.
